### 2019-06-10 thesis-template/msc-thesis.tex
update

@@

% Choose language
\newcommand{\setlang}[1]{\selectlanguage{#1}\nonfrenchspacing}
\newcommand*\mean[1]{\bar{#1}}
\begin{document}

@@

\chapter*{Abstract}
%\input{abstract}
Tensor fields, are most commonly associated with stress distributions (cf. Cauchy-stress tensor) in 2D/3D grids, but also have some other meanings in practice in physics. By means of global illumination techniques, we motivate a new method to visualize tensor fields, which is also capable of visualizing the light transport (propagation) in a topologically defined scene geometry. As a basis, we derive a simple light propagation scheme for Cartesian grids, which satisfies propagation attenuation and energy-conservation principles and is able to approximate light distributions for given light source position(s) and direction(s) until convergence. The transmission profiles within this grid are considered as crystal fiber structures as an underlying physical model (eg. gemstones: tiger's eye's cat's eye effect). The consequent task is to model anisotropic fibre structures with the orientation and anisotropy measures of the underlying tensor field in the grid.
For this, we derive an eigenbasis by PCA (principal component analysis) for every cell of the tensor field and form an ellipsoid equation as transmission (transfer/weighting) function for the transmitted light profiles. Hence, the resulting light distribution can be visualized via 2D scalar field or polar plots for directional information. Additionally, We measure impulse responses of the tensor field with Dirac-pulses as light sources at any (sampled) position and in any direction to generate a uniform sampled map of the global illumination distributions. This map is considered as a global illumination energy flow field.
We gain inspiration by vector field particle tracing's FTLE approach for analyzing the gradient of the flow field of the resulting light distributions. Similar to how the FTLE detects ridges/LCS (Lagrangian coherent structures) in vector fields (flow fields), our approach captures these structures likewise, attracting/repelling/saddling tensor field lines (TFL) which frequently represent key structures in nature, in tensor fields. We denote this entity a global illumination gradient which states an FTLE-related approach for visualizing LCS in tensor fields through analysis of the global and light transport  directed by imprint/engraving in crystelline fiber structures modeled by the tensor field's ellipsoid glyhps. We evaluate our approaches for plausibility and comparability within an extensive test campaign and demonstrate its performance on several examples of datasets.
\newpage

@@

In this work, we will discuss the advantages of tensor field analysis and motivate a new method for tensor field through a coherent light transport visualization. At first, we will shortly state the problem setting objectives and then give a quick overview about the structure of this work.

\paragraph{Contributions of this Work}
\begin{itemize}
	\item Motivation of a light transport technique following basic but crucial physical principles for tensor field visualization
	\item Design of a FTLE-related approach for visualizing key structures (ROIs) in $2D$ second-order tensor fields
	\item Discussion of the proposed techniques in the context of other works in this field
\end{itemize}
%Dieses Kapitel gibt einen Überblick über die Arbeit. Gerade der
%Abschnitt zur Motivation soll allgemein verständlich geschrieben
%werden. Die Einleitung sollte auch wichtige Referenzen enthalten.

\section{Motivation}
\label{sect:motiv}
Tensor Field Visualization is gaining importance as a relevant tool for the analysis of fluid and solid mechanics. Tensors are found most commonly in medical, scientific and engineering applications. A tensor is a $n$-D generalization of a matrix and appears, e.g., as Jacobian-Matrix in vector (Flow) field visualization or as stress tensor in solid continuum mechanics. Basically, tensors describe compressions, tensions, rotations and volume changes in both solid and fluid material. While most techniques focus on symmetric tensor field visualization (like glyphs and tensor field lines), some other recent works adress the problem of asymmetric tensor field visualization. The issue with asymmetric or antisymmetric tensors is that they do not always yield real eigenvalues, which is needed to determine an eigenbasis to set up glyphs, hyperstreamlines or tensor field lines. \cite{pang&zheng} Zheng \& Phang et al. proposed the concept of dual eigenvectors and \cite{zhang} Lin et al. extended it to pseudo-eigenvectors and introduced eigenvector and eigenvalue manifold to visualize eigenvectors in the complex domain. In this work, we will use the $\mathbf{U}$-matrix obtained from singular value decomposition (SVD) to obtain the half-axes of the PCA principal ellipsoid as suggested by Moler \cite{moler}.\\

\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/DIFFERENT_TYPES_OF_STRESS.png}
  }
  \caption{Different types of stress}
  \label{Skalierung}
\end{figure}

Worum geht es? Beispiel(e)! Illustrationen sind hier meist sinnvoll
zum Verständnis. Warum ist das Thema wichtig? In welchem Kontext?

\section{Objectives}
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/hyperstreamlines.png}
  }
  \caption{Hyperstreamlines in Tensor Field Visualization}
  \label{Skalierung}
\end{figure}
The main purpose of visualization in general is to reduce or encode some set of numbers in some readable, explorable and more intuitive representation. This implies, e.g., clustering, structuring or projection techniques. When it comes to tensor fields, this comes handy in, e.g., medical sciences such as in DT-MRI (diffusion tensor-medical resonance imaging) or in simulated (or measured) stress tensor fields from solid and fluid mechanics. The problem setting encomprises the visualization of tensor fields, i.e., mostly 2D/3D-grids consisting of mainly second-order tensors (cf. Cauchy-Stress Tensor) with the additional possibility to model light transport transport in a geometrical scene. The objective of this work is to design a simple light propagation scheme following the most basic but crucial physical principles and laws. This propagation scheme can either be used to visualize the light transport paths (separatrices) in a scene defined by geometry. Or, on the other hand, be used to visualize the anisotropy characteristics of a given tensor field. Therefore the tensor field somehow needs to be pre-processed to generate a tensor-induced footprint map of transmission profiles. Hence however, the tensors need to be decomposed to obtain a kind of transmission profile. Another aim is to implement an approach borrowed from vector-field visualization. We construct an FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges/LCS (Lagrangian coherent structures). These structures can be classified (identified) as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature. LCS usually separate regions of different flow behavior, but in our case they separate regions/domains of varying directional distributions (w.r.t. direction/location) represented by tensors. The map responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other because of bidirectional stress directions and since we effectively measure the FTLE with both forward and backward integration time (stimulus directions). The real part of the eigenvalues could be exploited for sign change (contractions/tensions) which leads to the concept of deformation glyphs w. attached arrows pointing inward/outward for the two types of normal stresses. Practically, in its eigenbasis there are no shear stresses existent for the symmetric stress tensor in equilibrium, just the principal normal stresses. Thus, we can even define the full set of stresses with the glyphs which are derived from the singular value decomposition and signed with the real part of the eigenvalues (in corresponding absolute order). The main goal of this work is therefore the encoding of a tensor field into certain elaborate representations and the interpretation of meaningful results comparing the proposed techniques against each other. The designed approach could e.g. be applied in medicinal sciences, engineering and physics to reveal structures which are hard to grasp through interpreting simpler glyph and tensor field line visualizations.

@@

For the rest of this work, we will give a quick overview on the structure and outline. Next, in chap. \ref{chap:grundlagen}, we will capture important fundamentals like Tensors and PCA in the context of singular value decomposition (generalization of eigenvalue decomposition). Also we will discuss the most relevant related work in the domain of tensor field visualization. In chap. \ref{chap:main} we will propose the main contributions of this work and document the experimental evaluation in chap. \ref{chap:eval}. A compact Summary, Conclusion and Outlook will be given in chap. \ref{chap:concl}.\\

@@

In this chapter, we will capture the prerequisites necessary to understand the contributions of this work and will capture Tensors, PCA, GT approaches and Related Work in that sense.

@@

\section{Tensor Fields}
\label{sec:tensors}
Now what is a Tensor? A tensor is a $o$-dimensional generalization of a matrix as illustrated in Table \ref{tensor}.
\begin{table}[!b]
\centering
\caption{Tensor Shapes}
 \begin{tabular}{c|c|c|c|c|c|c}
\textbf{order} & 0 & 1 & 2 & 3 & ... & o \\ 
\hline 
\textbf{shape} & scalar & vector & matrix & ``3D-matrix'' & ... & $\mathop{o}$D-matrix\\ 
\end{tabular}
\label{tensor}
 \end{table} 
In addition to that, a tensor has as many indices as its order $o$ and their run length is as long as its embedded dimension $n$ (the dimensin of the row and column vectors), which is equal to the rank of the tensor for full rank (or in other words the number of elements n for a $n{\times}n$-matrix). Tensors follow certain transformation rules which are defined for covariant/contravariant tensors. They are most commonly obtained and defined from physical transformation equations. Hence, any $o$-D number array could be a tensor, but the definition holds only if the transformation rules apply to these $o$-D number arrays (in a sense of physical interpretation), which is mostly true for component-indexed $o$-D number entities following matrix multiplication and transformation rules found in physics and math. In fact, tensors themselves are interpreted as a transformation which reflects an incoming direction vector (measuring, e.g., the stress in that direction) to an outgoing stress vector.
 To be a bit more concrete, tensors represent stress distributions in solids and fluids describing strain and flow features as these entities aren't simple numbers or vectors. They occur in form of a tensor fields, and thus as entities or quantities in physics. That is, at each point in space (indexed by the grid) there is a whole distribution of, e.g., stresses (elastic and viscous) which needs to be characterized and represented by a single tensor.
 \\
 \paragraph{stress tensor}
 \begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/stress_tensor.png}
  }
  \caption{Cauchy-Stress Tensor}
  \label{cauchy}
\end{figure}

The Cauchy stress tensor, which is propably the classical example of a tensor, is depicted in Fig. \ref{cauchy} and consists of 3 stress vectors $\mathbf{T(e_i)}$ arranged in row-major order. It appears in similar form as viscous stress tensor in fluid mechanics:
\begin{align*}
\boldsymbol{\sigma}
 = \left[{\begin{matrix} \mathbf{T}^{(\mathbf{e}_1)} \\
\boldsymbol{T}^{(\mathbf{e}_2)} \\
\boldsymbol{T}^{(\mathbf{e}_3)} \\
\end{matrix}}\right] =
\left[{\begin{matrix}
\sigma _{xx} & \sigma _{xy} & \sigma _{xz} \\
\sigma _{yx} & \sigma _{yy} & \sigma _{yz} \\
\sigma _{zx} & \sigma _{zy} & \sigma _{zz} \\
\end{matrix}}\right]  = \left[{\begin{matrix}
\sigma _x & \tau _{xy} & \tau _{xz} \\
\tau _{yx} & \sigma _y & \tau _{yz} \\
\tau _{zx} & \tau _{zy} & \sigma _z \\
\end{matrix}}\right] =
\begin{bmatrix}
\sigma _{11} & \sigma _{12} & \sigma _{13} \\
\sigma _{21} & \sigma _{22} & \sigma _{23} \\
\sigma _{31} & \sigma _{32} & \sigma _{33} 
\end{bmatrix}.
\end{align*}
These 3 stress vectors represent the orientation and magnitude of total resulting stress at plane ${x,y,z}$ in direction ${x,y,z}$ ($3^2=9$ individual numbers). Thus, the stress tensor poses a full valid represenation of the stress distribution at any point in space. The stress tensor itself can be interpreted as a transformation which maps an incoming normal direction vector $\mathbf{n}$ a resulting stress vector $\mathbf{T^{(n)}} = \mathbf{n}\cdot\boldsymbol{\sigma}$. The stresses in normal directions (diagonal elements) are directly related to the pressure, which is always isotropic and hence non-directionally dependent. It is possible to find 2 principal stress directions in 2D through PCA (see sect. \ref{sec:pca} below), for which there are no shear stresses existent in equilibrium. These form the principal directions of deformation at a certain location and are sufficient to describe the resulting transformation/deformation behaviour. The resulting system of principal directions and absolute stresses is often called an ``eigensystem'' or ``eigenframe''. So to say, an eigensystem, consistent of the principal directions/axes, spans a principal ellipsoid depicted in Fig. \ref{HAT} in sect. \ref{sec:pca}.
Another interpretation for the principal component ellipsoid is that it is the result of the tensor transformation applied to the unit circle. It is therefore somehow representing the distortion of an absolutely isotropic object (unit circle) by whatever the affine transformation effects (e.g. stretching, shearing, mirroring, scaling, rotating).
\section{Principal Component Analysis}
\label{sec:pca}
Note that we denote a matrix with a second order tensor interchangeably for the sake of simplicity in the following. For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent. Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted. The anisotropy can be represented by, e.g., ellipsoid glyphs.

\begin{figure}[!t]
%\centering
  \begin{minipage}{0.5\textwidth}
    \includegraphics[width=\textwidth]{img/HAT.png}
    \label{HAT}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \includegraphics[width=\textwidth]{img/glyphs.png}
    \label{b)}
  \end{minipage}
\caption{a) 2D Principal component ellipsoid (eigenframe), b) 3D ellipsoid glyphs w. anisotropy coefficient (l:linear, p:planar, s:spherical)}
\label{gradient}
\end{figure}

Principal Component Analysis is an algorithm which can capture the principal components (directions) in stochastic data. But it works similarly on transformations to capture the principal directions from transformations like rotation and scaling. Now, principal component analysis can be done by eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for asymmetric matrices. For (skew-)symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$ (cf. \cite{kieburg}). Note that the eigenvectors match the singular vectors only for symmetric matrices, which does not mean that they do not form the half-axes of the principal component ellipsoid in other cases. Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling ($\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$). We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive). Also, the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig. \ref{HAT} for any kind of matrix as pointed out by Moler \cite{moler}. In addition, we also calculate the eigenvalues to exploit the sign of the real part of the eigenvalues (\cite{mroz}) ordered decreasingly by absolute value (corresponding to singular value order). The sign corresponds to tensile ($+$) and compressive ($-$) stress, which would allow us to visualize deformation glyphs with arrows pointing inwards for compressive and outwards for tensile stresses. We do this decomposition analysis once for each matrix and then store the precomputed results in a grid. The principal ellipsoid is also used as a transmission profile for the propagation scheme in symbolic form in a subequent step. In this work, we will use the singular vectors yielded from singular value decomposition to avoid complex values yielded from eigenvalue decomposition. This will allow us to set up glyphs representing the major/minor axis of the underlying transformation for every kind of (asymmetric/symmetric) tensor. We will then use global-illumination techniques on top to generate a flow map, which permits us to compute an FTLE-like field on the tensor fields.

\section{GT concepts / Approaches}
\label{sec:gt}
\paragraph{Glyphs and tensor field lines}
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.5\textwidth]
    {img/tensorfieldlines.png}
  }
  \caption{white: bold lines; ellipsoids: ellipsoid glyphs}
  \label{scheme}
\end{figure}\vskip 3pt
Glyphs were derived as a basis for the weighting profiles of the tensors in the following propagation scheme. They were chosen because it is an easy-to-use (understand) concept which can also be interpreted nicely. In fact, these are used for the principal component analysis of direction vectors for the Tensor Field Line extrapolation (integration) in a subsequent step.
Tensor Field Lines (the type of tensor stream line - hyperstreamline analog that does not introduce artificial intertia) were implemented for Ground Truth (GT) data generation purposes. For example, when we design an FTLE-like field computation approach, we will need to know when tensor field lines diverge/converge.

@@

This chapter will summarize the main contributions of this work studied and state targeted aims and goals as a first step in sect. \ref{sec:overview}. Then we will have a short introduction to the implemented global illumination scheme derived and simplified from an nVidia GI approach in sect. \ref{sec:scheme}. Last, we will state the effectively implemented Visualization methods Light Transport Visualization and Global Illumination gradient proposed as innovatively invented tensor visualization techniques.

>>>>

\usepackage{caption}

@@

% Choose language
\newcommand{\setlang}[1]{\selectlanguage{#1}\nonfrenchspacing}
\newcommand*\mean[1]{\bar{#1}}
\newcommand{\source}[1]{\vspace{-3pt} \caption*{ Source: {#1}} }

\begin{document}

@@

\chapter*{Abstract}
%\input{abstract}
Many problems in science and engineering require
tensor field representations. Tensor fields, are most commonly associated with stress distributions (cf. Cauchy-stress tensor) in 2D/3D grids, but also have some other meanings in practice in physics. By means of global illumination techniques, we motivate a new method to visualize tensor fields, which is also capable of visualizing the light transport (propagation) in a topologically defined scene geometry. As a basis, we derive a simple light propagation scheme for Cartesian grids, which satisfies propagation attenuation and energy-conservation principles and is able to approximate light distributions for given light source position(s) and direction(s) until convergence. The transmission profiles within this grid are considered as crystal fiber structures as an underlying physical model (eg. gemstones: tiger's eye's cat's eye effect). The consequent task is to model anisotropic fibre structures with the orientation and anisotropy measures of the underlying tensor field in the grid.
For this, we derive an eigenbasis by PCA (principal component analysis) for every cell of the tensor field and form an ellipsoid equation as transmission (transfer/weighting) function for the transmitted light profiles. Hence, the resulting light distribution can be visualized via 2D scalar field or polar plots for directional information. Additionally, We measure impulse responses of the tensor field with Dirac-pulses as light sources at any (sampled) position and in any direction to generate a uniform sampled map of the global illumination distributions. This map is considered as a global illumination energy flow field.
We gain inspiration by vector field particle tracing's FTLE approach for analyzing the gradient of the flow field of the resulting light distributions. Similar to how the FTLE detects ridges/LCS (Lagrangian coherent structures) in vector fields (flow fields), our approach captures these structures likewise, attracting/repelling/saddling tensor field lines (TFL) which frequently represent key structures in nature, in tensor fields. We denote this entity a global illumination gradient which states an FTLE-related approach for visualizing LCS in tensor fields through analysis of the global and light transport  directed by imprint/engraving in crystelline fiber structures modeled by the tensor field's ellipsoid glyhps. We evaluate our approaches for plausibility and comparability within an extensive test campaign and demonstrate its performance on several examples of datasets.
\newpage

@@

In this work, we will discuss the advantages of tensor field analysis and motivate a new method for tensor field visualization through a coherent light transport visualization. At first, we will shortly state the problem setting, objectives and give a quick overview about the structure of this work. The main contributions of this work are listed in the bullet list below:

\paragraph{Contributions of this Work}
\begin{itemize}
	\item Motivation of a light transport technique following basic but crucial physical principles for tensor field visualization using tensors as transmission profiles
	\item Design of a FTLE-related approach called light transport gradient (LTG) for visualizing key structures (ROIs) in $2D$ second-order tensor fields
	\item Discussion of the proposed methods in the context of other works in this field
\end{itemize}
%Dieses Kapitel gibt einen Überblick über die Arbeit. Gerade der
%Abschnitt zur Motivation soll allgemein verständlich geschrieben
%werden. Die Einleitung sollte auch wichtige Referenzen enthalten.

\section{Motivation}
\label{sect:motiv}
Tensor field visualization is gaining importance as a relevant tool for the analysis of fluid and solid mechanics. Tensors are found most commonly in medical, scientific and engineering applications. In general, a tensor is a $o$-D generalization of a matrix and appears, e.g., as Jacobian matrix in vector (flow) field visualization or as stress tensor in solid continuum mechanics. Basically, tensors describe compressions, tensions, rotations and volume changes in both solid and fluid material. While most techniques focus on symmetric tensor field visualization (like glyphs and tensor field lines), some other recent works adress the problem of asymmetric tensor field visualization. The issue with asymmetric (non-symmetric) or antisymmetric (skew-symmetric) tensors is, that they do not always yield real eigenvalues, which is needed to determine an eigenbasis to set up glyphs, hyperstreamlines or tensor field lines. Nevertheless, there has been extensive work in this domain as well. Zheng \& Phang et al. \cite{pang&zheng} proposed the concept of dual eigenvectors and Lin et al. \cite{zhang} extended it to pseudo-eigenvectors and introduced eigenvector and eigenvalue manifold to visualize eigenvectors in the complex domain. In this work, we will use the $\mathbf{U}$-matrix obtained from singular value decomposition (SVD) to obtain the half-axes of the PCA principal ellipsoid representing the tensor transformation as suggested by Moler \cite{moler} to state another alternative of handling asymmetric tensors.\\

\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/DIFFERENT_TYPES_OF_STRESS.png}
  }
  \caption{Different types of stress}
  \captionsetup{font={footnotesize,bf,it}}
  \caption*{src: \url{https://upload.wikimedia.org/wikipedia/commons/e/eb/DIFFERENT_TYPES_OF_STRESS.png}}	
  \label{Skalierung}
\end{figure}

\section{Objectives}
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/hyperstreamlines.png}
  }
  \caption{Hyperstreamlines in tensor field visualization}
  \captionsetup{font={footnotesize,bf,it}}
  \caption*{src: \url{https://www-apr.lip6.fr/~tierny/stuff/teaching/tierny_visualization_tensorFields.pdf}}	
  \label{Skalierung}
\end{figure}
The main purpose of visualization in general is to reduce or encode some set of numbers in some readable, explorable and more intuitive representation. This implies, e.g., clustering, structuring or projection techniques. When it comes to tensor fields, this comes handy in, e.g., medical sciences such as in DT-MRI (diffusion tensor-medical resonance imaging) or in simulated (or measured) stress tensor fields from solid and fluid mechanics. The problem setting encomprises the visualization of tensor fields, i.e., mostly 2D/3D-grids consisting of mainly second-order tensors (cf. Cauchy stress tensor) with the additional possibility to model light transport transport in a geometrical scene. The objective of this work is to design a simple light propagation scheme yet following the most basic but crucial physical principles and laws, like energy conservation and propagation attenuation. This propagation scheme can either be used to visualize the light transport paths (defined by separatrices) in a scene defined by geometry. Or, on the other hand, be used to visualize the anisotropy characteristics of a given tensor field. However, the tensor field somehow needs to be pre-processed to generate a tensor-induced footprint map of transmission profiles. Hence however, the tensors need to be decomposed to obtain a kind of transmission profile. Another aim is to implement an approach borrowed from vector-field visualization. We construct an FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges/LCS (Lagrangian coherent structures). These structures can be classified (identified) as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature. LCS usually separate regions of different flow behavior, but in our case they separate regions/domains of varying directional distributions (w.r.t. direction/location) represented by tensors. The map responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other because of bidirectional stress directions and since we effectively measure the FTLE with both forward and backward integration time (stimulus directions). The real part of the eigenvalues could be exploited for sign change (contractions/tensions) which leads to the concept of deformation glyphs w. attached arrows pointing inward/outward for the two types of normal stresses. Practically, in its eigenbasis there are no shear stresses existent for the symmetric stress tensor in equilibrium, just the principal normal stresses. Thus, we can even define the full set of stresses with the singular values which are derived from singular value decomposition and signed with the real part of the eigenvalues (in corresponding absolute order). The main goal of this work is therefore the encoding of a tensor field into certain elaborate representations and the interpretation of meaningful results comparing the proposed techniques against others. The designed approach could e.g. be applied in medicinal sciences, engineering and physics to reveal structures which are hard to grasp through interpreting simpler and sometimes confusing glyph and tensor field line visualizations.

@@

For the rest of this work, we will give a quick overview on the structure and outline. Next, in chap. \ref{chap:grundlagen}, we will capture important fundamentals like Tensors, PCA and GT data in the context of singular value decomposition (generalization of eigenvalue decomposition). Also we will discuss the most relevant related work in the domain of tensor field visualization. In chap. \ref{chap:main}, we will propose the main contributions of this work and document the results and evaluation in chap. \ref{chap:eval}. Subsequently, A compact Summary, Conclusion and Outlook will be given in chap. \ref{chap:concl}.\\

@@

In this chapter, we will capture the prerequisites necessary to understand the contributions of this work and we will capture tensor fields, PCA, GT approaches and Related Work in that sense.

@@

\section{Tensor Fields}
\label{sec:tensors}
Now what is a tensor? A tensor is a $o$-dimensional generalization of a matrix as illustrated in Table \ref{tensor}.
\begin{table}[!b]
\centering
\caption{Tensor Shapes}
 \begin{tabular}{c|c|c|c|c|c|c}
\textbf{order} & 0 & 1 & 2 & 3 & ... & o \\ 
\hline 
\textbf{shape} & scalar & vector & matrix & ``3D matrix'' & ... & $\mathop{o}$D matrix\\ 
\end{tabular}
\label{tensor}
 \end{table} 
In addition to that, a tensor has as many indices as its order $o$ and their run length is as long as its embedded dimension $n$ (the dimensin of the row and column vectors), which is equal to the rank of the tensor for full rank (or in other words the number of elements n for a $n{\times}n$-matrix). Tensors follow certain transformation rules which are defined for covariant/contravariant tensors. They are most commonly obtained and defined from physical transformation equations. Hence, any $o$D number array could be a tensor, but the definition holds only if the transformation rules apply to these $o$D number arrays (in a sense of mathematical/formal interpretation), which is mostly true for component-indexed $o$-D number entities following matrix multiplication and transformation rules found in physics and math manifesting in calculation rules like matrix multiplication. Thus, a tensor is a $o$D number array arranged in meaningful order to put it simple. In fact, tensors themselves are interpreted as transformations which reflect an incoming direction vector (measuring, e.g., the stress in that direction) to an outgoing (stress) vector. To be a bit more concrete, tensors represent stress distributions in solids and fluids describing strain and flow features as these entities are not simple numbers or vectors. They occur in form of a tensor fields, and thus as entities or quantities in physics. That is, at each point in space (indexed by the grid) there is a whole distribution of directions, e.g., stresses (elastic and viscous) which needs to be characterized and represented by a single tensor each.
 \\
 \paragraph{stress tensor}
 \begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/stress_tensor.png}
  }
  \caption{Cauchy stress Tensor}
  \captionsetup{font={footnotesize,bf,it}}
  \caption*{src: \url{https://en.wikipedia.org/wiki/Cauchy_stress_tensor\#/media/File:Components_stress_tensor_cartesian.svg}}
  \label{cauchy}
\end{figure}

The Cauchy stress tensor, which is propably the classical example of a tensor, is depicted in Fig. \ref{cauchy} and consists of 3 stress vectors $\mathbf{T(e_i)}$ arranged in row-major order. It appears in similar form as viscous stress tensor in fluid mechanics:
\begin{align*}
\boldsymbol{\sigma}
 = \left[{\begin{matrix} \mathbf{T}^{(\mathbf{e}_1)} \\
\boldsymbol{T}^{(\mathbf{e}_2)} \\
\boldsymbol{T}^{(\mathbf{e}_3)} \\
\end{matrix}}\right] =
\left[{\begin{matrix}
\sigma _{xx} & \sigma _{xy} & \sigma _{xz} \\
\sigma _{yx} & \sigma _{yy} & \sigma _{yz} \\
\sigma _{zx} & \sigma _{zy} & \sigma _{zz} \\
\end{matrix}}\right]  = \left[{\begin{matrix}
\sigma _x & \tau _{xy} & \tau _{xz} \\
\tau _{yx} & \sigma _y & \tau _{yz} \\
\tau _{zx} & \tau _{zy} & \sigma _z \\
\end{matrix}}\right] =
\begin{bmatrix}
\sigma _{11} & \sigma _{12} & \sigma _{13} \\
\sigma _{21} & \sigma _{22} & \sigma _{23} \\
\sigma _{31} & \sigma _{32} & \sigma _{33} 
\end{bmatrix}.
\end{align*}
These 3 stress vectors represent the orientation and magnitude of total resulting stress at plane ${x,y,z}$ in direction ${x,y,z}$ ($3^2=9$ individual numbers). Thus, the stress tensor poses a full valid represenation of the stress distribution at any point in space. The stress tensor itself can be interpreted as a transformation which maps an incoming normal direction vector $\mathbf{n}$ a resulting stress vector $\mathbf{T^{(n)}} = \mathbf{n}\cdot\boldsymbol{\sigma}$. The stresses in normal directions (diagonal elements) are directly related to the pressure, which is always isotropic and hence non-directionally dependent (kind of a Euclidean distance of the stress components). It is possible to find 2 principal stress directions in 2D through PCA (see sect. \ref{sec:pca} below), for which there are no shear stresses existent in equilibrium. These form the principal directions of deformation at a certain location and are sufficient to describe the resulting transformation/deformation behaviour. The resulting system of principal directions and absolute stresses is often called an ``eigensystem'' or ``eigenframe''. So to say, an eigensystem, consistent of the principal directions/axes, spans a principal ellipsoid depicted in Fig. \ref{HAT} in sect. \ref{sec:pca}.
Another interpretation for the principal component ellipsoid is that it is the result of the tensor transformation applied to the unit circle. It is therefore somehow representing the distortion of an absolutely isotropic object (unit circle) by whatever the affine tensor transformation effects (e.g. stretching, shearing, mirroring, scaling, rotating).
\section{Principal Component Analysis}
\label{sec:pca}
Note that we denote a matrix with a second order tensor interchangeably for the sake of simplicity in the following. For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent. Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted. The anisotropy can be represented by, e.g., ellipsoid glyphs.

\begin{figure}[!t]
\hskip 22pt
  \begin{minipage}[!t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/HAT.png}
    \label{HAT}
   \caption{2D Principal component ellipsoid (eigenframe)}
   \captionsetup{font={footnotesize,bf,it}}
   \caption*{src: \url{https://de.wikipedia.org/wiki/Datei:HAT-ellipse0.svg}}
  \end{minipage} \hskip 40pt
  %\raggedright
  \begin{minipage}[!t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/glyphs.png}
    \caption{3D ellipsoid glyphs w. anisotropy coefficient (l:linear, p:planar, s:spherical)}
    \label{3Dglyphs}
    \captionsetup{font={footnotesize,bf,it}}
   \caption*{src: \url{https://slideplayer.com/slide/5291764/}}
  \end{minipage}
\label{gradient}
\end{figure}

Principal Component Analysis is an algorithm which can capture the principal components (directions) in stochastic data. But it works similarly on transformations to capture the principal directions from transformations like rotation and scaling. Now, principal component analysis can be done by eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for asymmetric matrices. For (skew-) symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$ (cf. \cite{kieburg}). Note that the eigenvectors match the singular vectors only for symmetric matrices, which does not mean that they do not form the half-axes of the principal component ellipsoid in other cases. Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling ($\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$). We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive). Also, the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig. \ref{HAT} for any kind of matrix as pointed out by Moler \cite{moler}. In addition, we also calculate the eigenvalues to exploit the sign of the real part of the eigenvalues (\cite{mroz}) ordered decreasingly by absolute value (corresponding to singular value order). The sign corresponds to tensile ($+$) and compressive ($-$) stress, which would allow us to visualize deformation glyphs with arrows pointing inwards for compressive and outwards for tensile stresses. We do this decomposition analysis once for each matrix and then store the precomputed results in a grid. The principal ellipsoid is also used as a transmission profile for the propagation scheme in symbolic form in a subequent step. In this work, we will use the singular vectors yielded from singular value decomposition to avoid complex values yielded from eigenvalue decomposition. This will allow us to set up glyphs representing the major/minor axis of the underlying transformation for every kind of (asymmetric/symmetric) tensor. We will then use global-illumination techniques on top to generate a flow map, which permits us somehow to direct the intensities which are given as pre-defined initialization values. This kind of light transport propagation scheme will also allow us to compute an FTLE-like field on the tensor fields.

\section{GT data / Approaches}
\label{sec:gt}
\paragraph{Glyphs and tensor field lines}
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.5\textwidth]
    {img/tensorfieldlines.png}
  }
  \caption{white: bold lines; ellipsoids: ellipsoid glyphs}
  \label{scheme}
\end{figure}\vskip 3pt
Glyphs were derived as a basis for the weighting profiles of the tensors in the following propagation scheme. They were chosen because it is an easy-to-use (understand) concept which can also be interpreted nicely. In fact, these are used for the principal component analysis of direction vectors for the Tensor Field Line extrapolation (integration) in a subsequent step.
Tensor Field Lines (the type of tensor stream line - hyperstreamline analog that does not introduce artificial intertia) were implemented for Ground Truth (GT) data generation purposes. For example, when we design an FTLE-like field computation approach, we will need to know when tensor field lines diverge/converge.

@@

This chapter will summarize the main contributions of this work studied and state targeted aims and goals as a first step in sect. \ref{sec:overview}. Then we will have a short introduction to the implemented global illumination scheme derived and simplified from an nVidia global illumination approach in sect. \ref{sec:scheme}. Last, we will state the effectively implemented Visualization methods Light Transport Visualization and Global Illumination gradient proposed as innovatively invented tensor visualization techniques.
